---
title: "2_SDMs_nightingales"
author: "Aimara Planillo"
date: "11/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SDM Nightingales Berlin

Different species distribution models based on the three datasets for Nightingales
in Berlin

- JSDM data (standardized data)
- eBird (semi-structured data)
- Forschungsfall Nachtigall Project (opportunistic unstructured data)


```{r packages}
source("./CS_SDM_NightingaleBerlin/source_packages.R")
```

```{r define workspace}
Work_dir <- getwd()

rawdata_wd <- file.path(Work_dir, "/2_Raw_data/BreedingMonitoringData2017")
procdata_wd <- file.path(Work_dir, "/4_Processed_data")
output_wd <- file.path(Work_dir, "5_Results/BreedingMonitoring2017")
```


```{r general spatial data}
# spatial data 
berlin <- read_sf("2_Raw_data/EnvCov/berlin_city_border_25833.gpkg", crs = 25833)
water <- read_sf("2_Raw_data/EnvCov/waterbodies_Berlin_25833.gpkg", crs = 25833)
# env variables
env_stack <- stack(paste0(procdata_wd, "/ForPrediction/env_cov_berlin_100m_25833.tif"))
```


########################################
## Standardized data - Binomial model ##
########################################

```{r load standardized observations}
# standardized monitoring observations
NG2017 <- read.csv(paste0(rawdata_wd,
                          "/Nightingales2017_BreedingBirdSurvey.csv"))
# monitoring transect variabels
transects <- read.csv(paste0(rawdata_wd,
                             "/Breeding_transects_variables.csv"))
# env variables for monitoring transects
for_predict_cov <- read.csv(paste0(rawdata_wd,
                             "/grid_topredict_100mres_birds_noNA.csv"))

All_data <- cbind(transects, NG2017[,2:3])
All_data <- All_data %>%
  mutate(Lm_presence = ifelse(Lm_counts >0, 1,0))
All_data

colnames(All_data)
predict_cov <- for_predict_cov %>%
  rename(tree = tree.100m, 
         o.green = open.green100m,
         noise = noise.100m,
         pop = pop.100m, 
         prey = prey.abu) %>%
  select(x, y, tree, o.green, noise, pop, prey)
head(predict_cov)

```

```{r binomial glm standardized data}
ng_binomial <- glm(Lm_presence ~ (scale(tree) + scale(o.green) + 
                                      scale(noise) + scale(pop)) * 
                       scale(prey),
                   data = All_data, 
                   family = "binomial")
summary(ng_binomial)

plot(allEffects(ng_binomial))

# Evaluate model with AUC
pred_binomial <- predict(ng_binomial, type = "response")
All_data$predicts <- pred_binomial

g <- roc(Lm_presence ~ predicts, data = All_data)
plot(g)

#create new covariate values within observed range for predictions
pred1 <- data.frame(o.green = seq(from = 0,
                                    to = 1, by = 0.05), 
                    tree = mean(All_data$tree), 
                    noise = mean(All_data$noise), 
                    pop = mean(All_data$pop),
                    prey = mean(All_data$prey))

Pred <- predict(ng_binomial, newdata = pred1, type = "response")

plot(x = All_data$o.green, y = All_data$Lm_presence)
lines(pred1$o.green, Pred)

## predict binomial model for Berlin
pred_binomial <- predict(ng_binomial, newdata = predict_cov, type = "response", re.form = ~0)

predict_cov$predictions <- pred_binomial
head(predict_cov, 10)
```

```{r put binomial model predictions in a raster}
# use env raster as template
tree_raster <- env_stack$env_cov_berlin_100m_25833.2
prediction_raster <- raster(tree_raster)

# extract the cell numbers that correspond with the cells with predicitons 
# based on coordinates
cells <- cellFromXY(prediction_raster, as.matrix(predict_cov[, c("x", "y")]))

# Assign values to raster cells
prediction_raster[cells] <- predict_cov$predictions
prediction_raster <- mask(prediction_raster, berlin)
plot(prediction_raster)

writeRaster(prediction_raster, paste0(output_wd, "/raster_binomial_prediction_100m_25833.tif"), 
            overwrite = TRUE)

#Plot predictions 
tmap_mode("plot")
tm_shape(prediction_raster) +
  tm_raster(palette = rev(viridis(10)), 
            title = "Binomial prediction")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
   tm_layout(main.title = "Binomial GLM prediction Nightingales 2017\nselected covariates", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("left", "bottom"),
             size = 1.5) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 0.6) +
  tm_credits(paste0("AUC = ", round(g$auc[1],3)), position=c("left", "top"))

tmap_save(filename = paste0(output_wd, "/Predictions_binomial_2017.png"))
```



#################################################
## Semi-structured data - eBird Binomial model ##
#################################################

```{r load and prepare ebird data}
# species observation data
eBird_data <- read.csv(paste0(procdata_wd,
                          "/eBird/ebd_marjul201718_nightingale_zf_ss1km.csv"))
table(eBird_data$species_observed, year(eBird_data$observation_date))

eBird_data_proj <- st_as_sf(eBird_data, coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 25833) 

# extract coordinates cells 100m ebirds with observations
r_tree <- raster(paste0(rawdata_wd, "/EnvCov/tree_cover_20m_25833.tif")) 
plot(r_tree)

r_agg <- aggregate(r_tree, fact = 5)

plot(r_agg)
plot(st_geometry(eBird_data_proj), add = TRUE)

p <- as(r_agg, 'SpatialPixels')
p1 <- as.data.frame(p)

# get cells form projected nightingales observations
cells.ebird <- cellFromXY(r_agg, st_coordinates(eBird_data_proj)[, c("X", "Y")]) 
cells.envcov <- cellFromXY(r_agg, as.matrix(envcov)[, c("x", "y")])

envcov_cells <- cbind(envcov, cells.envcov)

All_data_ebird <- cbind(eBird_data, cells.ebird)
```

```{r run binomial model for semi-structured ebird data}
data_ebird_models_2018 <- All_data_ebird %>%
  mutate(Lm_presence = ifelse(species_observed == TRUE, 1,0)) %>%
  mutate(year = year(observation_date)) %>%
  filter(year == 2018) %>%
  filter(protocol_type %in% c("Stationary", "Traveling")) %>%
  left_join(envcov_cells, by = c("cells.ebird" = "cells.envcov"))
  
table(data_ebird_models_2018$Lm_presence)

colnames(data_ebird_models_2018)

# Binomial GLM include effort covariates 2018
data_ebird_models_2018 <- data_ebird_models_2018 %>%
  mutate(day = yday(observation_date)) 

colnames(data_ebird_models_2018)

ng_binomial_effort_2018 <- glm(Lm_presence ~ (scale(tree.100m) + scale(open.green100m) + 
                                      scale(noise.100m) + scale(pop.100m)) * 
                          scale(prey.abu) + 
                          day + time_observations_started + duration_minutes + number_observers +
                          protocol_type + effort_distance_km,
                   data = data_ebird_models_2018, 
                   family = "binomial")
summary(ng_binomial_effort_2018)

plot(allEffects(ng_binomial_effort_2018))

testDispersion(ng_binomial_effort_2018)
modeleffort2018_Output <- simulateResiduals(fittedModel = ng_binomial_effort_2018, plot = T)
testResiduals(modeleffort2018_Output)

# Evaluate model effort 2018 with AUC
pred_binomial_effort2018 <- predict(ng_binomial_effort_2018, type = "response")
data_ebird_models_2018$predicts_effort <- pred_binomial_effort2018

g_effort2018 <- roc(Lm_presence ~ predicts_effort, data = data_ebird_models_2018)
g_effort2018
plot(g_effort2018)

#create new covariate values within observed range for predictions
# example open green area
# to get the hour with max detections
summary(data_ebird_models_2018$time_observations_started)
hist(data_ebird_models_2018$time_observations_started, breaks = seq(4,22, by = 1))
# Higher freq at 9

pred_effort2018 <- data_ebird_models_2018 %>%
  mutate(open.green100m = seq(from = 0,
                                    to = 1, length.out = nrow(data_ebird_models_2018)), 
         tree.100m = mean(tree.100m),
         noise.100m = mean(noise.100m), 
         pop.100m = mean(pop.100m),
         prey.abu = mean(prey.abu), 
         day = median(day), 
         time_observations_started = 9, # higher observations
         duration_minutes = 60, # standardized to 1 hour
         number_observers = 1, # standardized to 1 observer
         protocol_type = "Traveling", # standardized to travelling protocol
         effort_distance_km = 1) # standardized to 1 km

Pred_effort <- predict(ng_binomial_effort_2018, newdata = pred_effort2018, type = "response")

plot(x = data_ebird_models_2018$open.green100m, y = data_ebird_models_2018$Lm_presence)
lines(pred_effort2018$open.green100m, Pred_effort)
```

```{r predict semi-structured ebird binomial model berlin}
pred_binomial2018 <- predict(ng_binomial_2018, newdata = envcov, type = "response", re.form = ~0)

envcov$predictions2018 <- pred_binomial2018
head(envcov, 10)

# add effort to environmental covariates to predict
envcov_effort2018 <- envcov %>%
  mutate(day = median(data_ebird_models_2018$day), 
         time_observations_started = 10, # higher observations
         duration_minutes = 60, # standardized to 1 hour
         number_observers = 1, # standardized to 1 observer
         protocol_type = "Traveling", # standardized to travelling protocol
         effort_distance_km = 1) # standardized to 1 km
  
pred_binomial_effort2018 <- predict(ng_binomial_effort_2018, newdata = envcov_effort2018, type = "response", re.form = ~0)

envcov_effort2018$predictions2018 <- pred_binomial_effort2018
head(envcov_effort2018, 10)
```

```{r plot prediction map semi-structured ebird binomial model}
# Assign values to raster cells
prediction_raster_effort2018 <- prediction_raster
prediction_raster_effort2018[cells] <- envcov_effort2018$predictions2018
prediction_raster_effort2018 <- mask(prediction_raster_effort2018, berlin)
plot(prediction_raster_effort2018)

writeRaster(prediction_raster_effort2018, paste0(results_wd, "/raster_binomial_prediction_effort_2018_100m_25833.tif"), 
            overwrite = TRUE)

# binomial model, with effort
tmap_mode("plot")

tm_shape(prediction_raster_effort2018) +
  tm_raster(palette = rev(viridis(10)), 
            title = "Binomial prediction with effort")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
   tm_layout(main.title = "Binomial GLM prediction Nightingales 2018\neBird data - selected covariates + effort", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("left", "bottom"),
             size = 1.5) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 0.6) +
  tm_credits(paste0("AUC = ", round(g_effort2018$auc[1],3)), position=c("left", "top"))

tmap_save(filename = paste0(results_wd, "/Predictions_binomial_effort_2018_ebird.png"))
```



###############################################
## Semi-structured data - ebird MaxEnt model ##
###############################################

```{r load data for ebird maxent}
eBird_maxent_data <- read.csv(paste0(procdata_wd, "/eBird/ebd_marjul_201718_nightingale_all_presences.csv"))

eBird_maxent_data_2018 <- eBird_maxent_data %>% 
  mutate(year = year(observation_date)) %>%
  filter(year == 2018) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 25833) %>%
  st_coordinates() %>%
  as.data.frame() 

# berlin environment
tree_100m <- raster(paste0(procdata_wd, "/ForPrediction/tree_100m.asc"), crs = CRS('+init=EPSG:25833'))
ogreen_100m <- raster(paste0(procdata_wd, "/ForPrediction/ogreen_100m.asc"), crs = CRS('+init=EPSG:25833'))
noise_100m <- raster(paste0(procdata_wd, "/ForPrediction/noise_100m.asc"), crs = CRS('+init=EPSG:25833'))
hpop_100m <- raster(paste0(procdata_wd, "/ForPrediction/hpop_100m.asc"), crs = CRS('+init=EPSG:25833'))
preyabu <- raster(paste0(procdata_wd, "/ForPrediction/preyabu_100m.asc"), crs = CRS('+init=EPSG:25833'))
env_cov <- stack(tree_100m, ogreen_100m, noise_100m, hpop_100m, preyabu)
names(env_cov) <- c("tree_100m", "ogreen_100m", "noise_100m", "hpop_100m", "preyabu_100m")
env_cov <- mask(env_cov, berlin)

plot(env_cov)

# divide in training and test data
set.seed(102) # set seed for reproducibility
group <- kfold(eBird_maxent_data_2018, 5)
pres_train2018 <- as.data.frame(eBird_maxent_data_2018[group != 1, ])
pres_test2018 <- as.data.frame(eBird_maxent_data_2018[group == 1, ])

# select random points
bg_2018 <- randomPoints(env_cov, 10000)

# divide in training and test background points
set.seed(102)
group <- kfold(bg, 5)
bg_train2018 <- as.data.frame(bg[group != 1, ])
bg_test2018 <- as.data.frame(bg[group == 1, ])

```

```{r run maxent model ebird}
# specifying background and presence points for training the model
maxent_ebird2018 <- maxent(x = env_cov, p = pres_train2018, a = bg_train2018,
                  removeDuplicates = TRUE, path = paste0(results_wd, "/Maxent2018"))

# evaluate model - training points
eval.maxent2018 <- evaluate(p = pres_train2018, a = bg_train2018, maxent_ebird2018, env_cov)
eval.maxent2018

# evaluate model - test points
eval.maxent2018_test <- evaluate(p = pres_test2018, a = bg_test2018, maxent_ebird2018, env_cov)
eval.maxent2018_test

# Covariate effects and importance
png(paste0(results_wd, "/Covariate_responses2018.png"))
response(maxent_ebird2018)
dev.off()

png(paste0(results_wd, "/Covariate_importance2018.png"))
plot(maxent_ebird2018)
dev.off()

var_imp <- plot(maxent_ebird2018)
write.csv(as.data.frame(var_imp), paste0(results_wd, "/Covariate_importance_maxent_2018.csv"))
```

```{r predict ebird maxent Berlin}
# predict model
pberlin.maxentebird2018 <- predict(env_cov, maxent_ebird2018, progress='')

png(paste0(results_wd, "/Prediction_berlin2018.png"), 
    width = 3000, height = 2000, pointsize = 30)

plot(pberlin.maxentebird2018, main='Maxent, raw values')
plot(st_geometry(berlin), add=TRUE, border='dark grey')
dev.off()

writeRaster(pberlin.maxentebird2018, paste0(results_wd, "/Maxent_ebird_2018_raster_100m_25833.tif"))
```

```{r plot map predictions ebird maxent Berlin}
tmap_mode("plot")
tm_shape(pberlin.maxentebird2018) +
  tm_raster(palette = rev(viridis(10)), 
            title = "Maxent, raw values")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
   tm_layout(main.title = "Maxent prediction Nightingales 2018\neBird - selected covariates", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("left", "bottom"),
             size = 1.5) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 0.6) +
  tm_credits(paste0("AUC (test)= ", round(eval.maxent2018_test@auc, 3),"\n no background reductions \nor bias\ntest points = 1/5"), position=c("left", "top"))

tmap_save(filename = paste0(results_wd, "/Predictions_maxent_nobg_nobias_2018.png"))
```



#########################################################
## Unstructured data - Nachtigall project MaxEnt model ##
#########################################################

```{r load unstructured Nachtigall data}
# species records
nightingales_2018 <- read.csv(paste0(procdata_wd, "/Nightingales_corrected_2018_maxent.csv"))
nightingales_2018map <- st_as_sf(nightingales_2018, coords = c("X", "Y"), crs = 25833)

# background environment
tree_bg <- raster(paste0(procdata_wd, "/Background1000m/tree_100m.asc"), crs = CRS('+init=EPSG:25833'))
ogreen_bg <-raster(paste0(procdata_wd, "/Background1000m/ogreen_100m.asc"), crs = CRS('+init=EPSG:25833'))
noise_bg <- raster(paste0(procdata_wd, "/Background1000m/noise_100m.asc"), crs = CRS('+init=EPSG:25833'))
hpop_bg <- raster(paste0(procdata_wd, "/Background1000m/hpop_100m.asc"), crs = CRS('+init=EPSG:25833'))
preyabu_bg <- raster(paste0(procdata_wd, "/Background1000m/preyabu_100m.asc"), crs = CRS('+init=EPSG:25833'))
env_cov_bg <- stack(tree_bg, ogreen_bg, noise_bg, hpop_bg, preyabu_bg)
names(env_cov_bg) <- c("tree_100m", "ogreen_100m", "noise_100m", "hpop_100m", "preyabu_100m")

# berlin environment
tree_100m <- raster(paste0(procdata_wd, "/ForPrediction/tree_100m.asc"), crs = CRS('+init=EPSG:25833'))
ogreen_100m <- raster(paste0(procdata_wd, "/ForPrediction/ogreen_100m.asc"), crs = CRS('+init=EPSG:25833'))
noise_100m <- raster(paste0(procdata_wd, "/ForPrediction/noise_100m.asc"), crs = CRS('+init=EPSG:25833'))
hpop_100m <- raster(paste0(procdata_wd, "/ForPrediction/hpop_100m.asc"), crs = CRS('+init=EPSG:25833'))
preyabu <- raster(paste0(procdata_wd, "/ForPrediction/preyabu_100m.asc"), crs = CRS('+init=EPSG:25833'))
env_cov <- stack(tree_100m, ogreen_100m, noise_100m, hpop_100m, preyabu)
names(env_cov) <- c("tree_100m", "ogreen_100m", "noise_100m", "hpop_100m", "preyabu_100m")
env_cov <- mask(env_cov, berlin)

plot(env_cov)
plot(env_cov_bg)

# background points
bg_points <- read.csv(paste0(procdata_wd, "/Background1000m/background_points_bias_1000m_25833.csv"))

# inspect result by plotting
plot(env_cov_bg$tree_100m, legend=FALSE)
plot(st_geometry(nightingales_2018map), add = TRUE, cex=0.3)
points(bg_points, cex=0.3, col = "red")

# divide background into test and training sets 
set.seed(102)
group1 <- kfold(bg_points, 5)
backg1_train <- bg_points[group1 != 1, ]
backg1_test <- bg_points[group1 == 1, ]

# divide observations into test and training sets
set.seed(102)
group <- kfold(nightingales_2018map, 5)
pres_train <- as.data.frame(st_coordinates(nightingales_2018map[group != 1, ]))
pres_test <- as.data.frame(st_coordinates(nightingales_2018map[group == 1, ]))

```

```{r run unstructured Nachtigall maxent model}
# specifying background and presence points for training the model
# Model with only linear to approximate glm models 

maxent_bias1 <- maxent(x = env_cov_bg, p = pres_train, a = backg1_train,
                  removeDuplicates = TRUE, path = results_wd, args = c("-P", "-J", "-q", "-t", "-h")) #P: remove variable effects, J: remove jacknife, q: quadratic features in the formula, h: hinge features in the formula
maxent_bias1

# evaluate model - training points
eval.maxentbias <- evaluate(maxent_bias1, p=pres_train, a=backg1_train, x=env_cov_bg)
eval.maxentbias

# evaluate model - test points
eval.maxentbias1 <- evaluate(maxent_bias1, p=pres_test, a=backg1_test, x=env_cov_bg)
eval.maxentbias1

plot(eval.maxentbias1, 'ROC')

# Covariate effects and importance
plot(maxent_bias1)

# effect on responses
png(paste0(results_wd, "/Covariate_responses.png"))
response(maxent_bias1)
dev.off()

png(paste0(results_wd, "/Covariate_importance.png"))
plot(maxent_bias1)
dev.off()

var_imp <- plot(maxent_bias1)
write.csv(as.data.frame(var_imp), paste0(results_wd, "/Covariate_importance.csv"))
```

```{r predict model in background area}
# predict model
px.maxentbias1 <- predict(env_cov_bg, maxent_bias1, progress='')

# Obtain threshold for presence/absence
tr <- threshold(eval.maxentbias1, 'spec_sens')


png(paste0(results_wd, "/Prediction_background.png"), 
    width = 3000, height = 2000, pointsize = 30)
par(mfrow=c(1,2))

plot(px.maxentbias1, main='Maxent, raw values')
plot(st_geometry(berlin), add=TRUE, border='dark grey')

plot(px.maxentbias1 > tr, main='presence/absence')
plot(st_geometry(berlin), add=TRUE, border='dark grey')
points(pres_train, pch='+')

dev.off()

par(mfrow=c(1,1))

tmap_mode("plot")
tm_shape(px.maxentbias1) +
  tm_raster(palette = "viridis", 
            title = "Maxent, raw values")

```

```{r predict unstructured Nachtigall maxent model in Berlin}
# predict model
pberlin.maxentbias1 <- predict(env_cov, maxent_bias1, progress='')

# Obtain threshold for presence/absence
tr <- threshold(eval.maxentbias1, 'spec_sens')

png(paste0(results_wd, "/Prediction_berlin.png"), 
    width = 3000, height = 2000, pointsize = 30)
par(mfrow=c(1,2))

plot(pberlin.maxentbias1, main='Maxent, raw values')
plot(st_geometry(berlin), add=TRUE, border='dark grey')

plot(pberlin.maxentbias1 > tr, main='presence/absence')
plot(st_geometry(berlin), add=TRUE, border='dark grey')
points(pres_train, pch='+')

dev.off()

par(mfrow=c(1,1))

writeRaster(pberlin.maxentbias1, paste0(results_wd, "/Maxent_berlin_raster_100m_25833.tif"))
```

```{r plot map predictions unstructured Nachtigall maxent}
tmap_mode("plot")
tm_shape(pberlin.maxentbias1) +
  tm_raster(palette = rev(viridis(10)), 
            title = "Maxent, raw values")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
   tm_layout(main.title = "Maxent prediction Nightingales 2018\nselected covariates, with bias", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("left", "bottom"),
             size = 1.5) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 0.6) +
  tm_credits(paste0("AUC (test)= ", round(eval.maxentbias1@auc, 3),"\nBG buffer: 1000m \nBG points: 10000 \ntest points = 1/5"), position=c("left", "top"))

tmap_save(filename = paste0(results_wd, "/Predictions_maxentbias1_2018.png"))
```



#########################
# COMPARING PREDICTIONS #
#########################

```{r load nightingale data}
# Nachtigall project records
nightingales_2018 <- read.csv(paste0(procdata_wd, "/Nachtigall_project/Nightingales_corrected_2018_maxent.csv")) %>%
  st_as_sf(coords = c("X", "Y"), crs = 25833)

# eBird records
eBird_presences_2018 <- read.csv(paste0(procdata_wd, "/eBird/ebd_marjul_201718_nightingale_all_presences.csv")) %>%
  mutate(year = year(observation_date)) %>%
  filter(year == 2018) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 25833)

# Standardized data records
nightingales_transects <- read.csv(paste0(rawdata_wd, "/BreedingBirdMonitoringData_2017/Nightingales2017_BreedingBirdSurvey.csv"))
transects <- read.csv(paste0(rawdata_wd, "/BreedingBirdMonitoringData_2017/Breeding_transects_variables.csv"))

nightingales_transects2 <- cbind.data.frame(nightingales_transects, transects)
nightingales_transects_map <- nightingales_transects2 %>%
  dplyr::select(transect, Lm_counts, Lm_km, X, Y) %>%
  mutate(Lm_presence = ifelse(Lm_counts>0, 1, 0)) %>%
  st_as_sf(coords = c("X", "Y"), crs = 25833)
nightingales_transects_map

nightingales_quadrants <- read_sf(paste0(rawdata_wd, "/EnvCov/birds_quadrants1x1km_3035.gpkg"))
plot(st_geometry(nightingales_quadrants))
```

```{r load raster predictions}
# Jsdm abundance maps
jsdm_raster <- raster("7_Results/Nightingales_JSDM2017_predictions_100m_25833.tif")
# transform jsdm raster into probabilities to compare with the other maps
jsdm_prob_raster <- jsdm_raster/cellStats(jsdm_raster, max)

# Binomial model - Breeding birds 
binomial_bb_raster <- raster("7_Results/final_results/breeding_birds_models/raster_binomial_prediction_100m_25833.tif")

# Binomial model - eBird 
binomial_ebird_raster <- raster("7_Results/final_results/ebird_models/raster_binomial_prediction_2018_100m_25833.tif")

# Maxent model - eBird 
maxent_ebird_raster <- raster("7_Results/final_results/ebird_models/Maxent_ebird_2018_raster_100m_25833.tif")

# Maxent model - Nachtigall 2018
maxent_nachtigall_raster <- raster( "7_Results/final_results/nachtigall_models/Maxent_berlin_raster_100m_25833.tif")
```


```{r load env variables}
# berlin context
berlin <- read_sf(paste0(rawdata_wd, "/EnvCov/berlin_city_border_25833.gpkg"), crs = 25833) %>%
  st_union()
water <- read_sf(paste0(rawdata_wd, "/EnvCov/waterbodies_Berlin_25833.gpkg"))
tree_raster <- raster(paste0(procdata_wd, "/ForPrediction/tree_100m.asc"))
crs(tree_raster) <- CRS("+init=epsg:25833")
land_use <- read_sf(paste0(procdata_wd, "/berlin_LandUse_modified_32633.gpkg"),
                    crs = 32633) %>% st_transform(crs = 25833)

# background polygons
background.pol <- read_sf(paste0(procdata_wd, "/Background1000m/background_polygon_1000m_25833.gpkg"), crs = 25833) %>%
  st_sf(st_union(background.pol))

# berlin environmental covariates
env_cov <- stack(paste0(procdata_wd, "/ForPrediction/env_cov_berlin_100m_25833.tif")) 
namefile2 <- read.csv(paste0(procdata_wd, "/ForPrediction/names_env_cov_berlin_100m_25833.csv"))
names(env_cov) <- namefile2$x

# background environmental covariates
env_cov_bg <- stack(paste0(procdata_wd, "/Background1000m/env_cov_berlin1000m_100m_25833.tif")) 
namefile <- read.csv(paste0(procdata_wd, "/Background1000m/names_env_cov_berlin1000m_100m_25833.csv"))
names(env_cov_bg) <- namefile$x
```



#~~~~~~~~~~~~~~~#
# Plot SDM maps #
#~~~~~~~~~~~~~~~#


```{r abundances JSDM prediction}
tmap_mode("plot")

# With observations in transects

tm_shape(jsdm_raster) +
  tm_raster(palette = rev(viridis(10)), 
            title = "JSDM predicted \nabundance",
            breaks = c(0,2,4,6), 
            style = "cont")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
  tm_shape(nightingales_transects_map) +
  tm_symbols(size = 1, col = "Lm_km", border.col = "black", border.lwd = 2, 
             palette = rev(viridis(10)), 
             legend.col.show = FALSE)+
  tm_add_legend(type = "symbol", labels = "monitoring transects", shape = 21, 
                size = 1, border.lwd = 2, border.col = "black", col = "white")+
   tm_layout(main.title = "JSDM Nightingales 2017\n abundances", 
            compass.type = "arrow",
            legend.title.size = 1,
            legend.text.size = 0.8, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 3) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 1)

tmap_save(filename = paste0(results_wd, "/JSDM_abundances_transects_map.png"), 
          width = 8, height = 7, units = "in")
```

```{r probabilities JSDM prediction}
# With observations in transects

tm_shape(jsdm_prob_raster) +
  tm_raster(palette = rev(viridis(10)), 
            title = "JSDM estimated \nprobabilities",
            breaks = c(0,0.2,0.4, 0.6, 0.8, 1))+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
  tm_shape(nightingales_transects_map) +
  tm_symbols(size = 1, col = "Lm_km", border.col = "black", border.lwd = 2,
             style = "cont", palette = rev(viridis(10)),
             title.col = "",
             legend.col.show = TRUE)+
  tm_layout(main.title = "JSDM Nightingales 2017\nprobabilities based on abundance", 
            compass.type = "arrow",
            legend.title.size = 1,
            legend.text.size = 0.8, 
            legend.bg.color = "white",
            legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
            legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 1.2) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 0.5) 
  
tmap_save(filename = paste0(results_wd, "/JSDM_probabilites_map.png"), 
          width = 4.33, height = 3.86, units = "in")
```

```{r plot binomial standardized monitoring predictions}
# With observations in transects
tm_shape(binomial_bb_raster) +
  tm_raster(palette = rev(viridis(10)), 
            title = "Binomial GLM \nprobabilities",
            breaks = c(0,0.2,0.4, 0.6, 0.8, 1))+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
  tm_add_legend(type = "symbol", labels = "Observed abundance", shape = 21, 
                size = 1, border.lwd = 2, border.col = "black", col = "white")+
  tm_shape(nightingales_transects_map) +
  tm_symbols(size = 1, col = "Lm_km", border.col = "black", border.lwd = 2,
             style = "cont", palette = rev(viridis(10)),
             title.col = "",
             legend.col.show = TRUE)+
  tm_layout(main.title = "Nightingales 2018\nMaxent probabilities selected covariates", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
            legend.bg.color = "white",
            legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
            legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 3) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 1) 

tmap_save(filename = paste0(results_wd, "/Binomial_GLM_breeding_birds_2017_transects_map.png"), 
          width = 8, height = 7, units = "in")
```

```{r plot binomial semi-structured ebird GLM predictions}
# With observations in transects

tm_shape(binomial_ebird_raster) +
  tm_raster(palette = rev(viridis(10)), 
            title = "Binomial GLM \nprobabilities",
            breaks = c(0,0.2,0.4, 0.6, 0.8, 1))+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
  tm_add_legend(type = "symbol", labels = "Observed abundance", shape = 21, 
                size = 1, border.lwd = 2, border.col = "black", col = "white")+
  tm_shape(nightingales_transects_map) +
  tm_symbols(size = 1, col = "Lm_km", border.col = "black", border.lwd = 2,
             style = "cont", palette = rev(viridis(10)),
             title.col = "",
             legend.col.show = TRUE)+
  tm_layout(main.title = "Nightingales 2018\nMaxent probabilities selected covariates", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
            legend.bg.color = "white",
            legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
            legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 3) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 1) 

tmap_save(filename = paste0(results_wd, "/Binomial_GLM_eBird_2018_map.png"), 
          width = 8, height = 7, units = "in")
```


```{r plot semi-structured ebird maxent predictions}
# With observations in transects
tm_shape(maxent_ebird_raster) +
  tm_raster(palette = rev(viridis(10)), 
            title = "Maxent model \nprobabilities",
            breaks = c(0,0.2,0.4, 0.6, 0.8, 1))+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
  tm_shape(nightingales_transects_map) +
  tm_symbols(size = 1, col = "Lm_km", border.col = "black", border.lwd = 2,
             style = "cont", palette = rev(viridis(10)),
             title.col = "",
             legend.col.show = TRUE)+
  tm_layout(main.title = "Nightingales 2018\nMaxent probabilities selected covariates", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
            legend.bg.color = "white",
            legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
            legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 1.2) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 0.5) 

tmap_save(filename = paste0(results_wd, "/Maxent_eBird_2018_selcov_map.png"), 
          width = 4.33, height = 3.86, units = "in")

```


```{r plot unstructured Nachtigall maxent predictions}
# With observations in transects
tm_shape(maxent_nachtigall_raster) +
  tm_raster(palette = rev(viridis(10)), 
            title = "Maxent model \nprobabilities",
            breaks = c(0,0.2,0.4, 0.6, 0.8, 1))+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
  tm_shape(nightingales_transects_map) +
  tm_symbols(size = 1, col = "Lm_km", border.col = "black", border.lwd = 2,
             style = "cont", palette = rev(viridis(10)),
             title.col = "",
             legend.col.show = TRUE)+
  tm_layout(main.title = "Nightingales 2018\nMaxent probabilities selected covariates", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
            legend.bg.color = "white",
            legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
            legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 1.2) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 0.5) 
  
tmap_save(filename = paste0(results_wd, "/Maxent_Nachtigall_2018_selcov_map.png"), 
          width = 4.33, height = 3.86, units = "in")
```


#~~~~~~~~~~~~~~~~~~#
# Sustracting maps #
#~~~~~~~~~~~~~~~~~~#

Lets compare binomial and maxent maps with the JSDM by sustracting both probabilistic maps

Using Jsdm as a reference, 
- values of **zero** will show areas with the **same prediction**,
- **positive** values will show areas where probability is **lower in maxent model**
- **negative** values will show areas where probability is **higher in maxent model**.

## Standardized monitoring binomial glm 
```{r}
jsdm_binomial_breedingbirds_raster <- jsdm_prob_raster - binomial_bb_raster

tm_shape(jsdm_binomial_breedingbirds_raster) +
  tm_raster(palette = "PuOr", 
            title = "Difference \nJSDM - Binomial GLM\nBreeding birds",
            breaks = c(-1,-0.5,0, 0.5, 1),
            style = "cont")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
   tm_layout(main.title = "Jsdm - binomial GLM - Breeding birds", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 3) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 1) 

tmap_save(filename = paste0(results_wd, "/Comparison_JSDM-BinomialGLM_breedingbirds_2017_map.png"), 
          width = 8, height = 7, units = "in")
```

## semi-structured ebird binomial glm 
```{r}
jsdm_binomial_ebird_raster <- jsdm_prob_raster - binomial_ebird_raster

tm_shape(jsdm_binomial_ebird_raster) +
  tm_raster(palette = "PuOr", 
            title = "Difference \nJSDM - Binomial GLM\neBird data",
            breaks = c(-1,-0.5,0, 0.5, 1),
            style = "cont")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
   tm_layout(main.title = "Jsdm - binomial GLM - eBird", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 3) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 1) 

tmap_save(filename = paste0(results_wd, "/Comparison_JSDM-BinomialGLM_ebird_2018_map.png"), 
          width = 8, height = 7, units = "in")
```


## Semi-structured Maxent ebird
```{r}
jsdm_maxent_ebird_raster <- jsdm_prob_raster - maxent_ebird_raster

tm_shape(jsdm_maxent_ebird_raster) +
  tm_raster(palette = "PuOr", 
            title = "Difference \nJSDM - Maxent \neBird data",
            breaks = c(-1,-0.5,0, 0.5, 1),
            style = "cont")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
   tm_layout(main.title = "Jsdm - Maxent model \nselected covariates", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 3) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 1) 

tmap_save(filename = paste0(results_wd, "/Comparison_JSDM-Maxent_ebird_map.png"), 
         width = 8, height = 7, units = "in")
```

## Unstructured Maxent nachtigall
```{r}
jsdm_maxent_nachtigall_raster <- jsdm_prob_raster - maxent_nachtigall_raster

tm_shape(jsdm_maxent_nachtigall_raster) +
  tm_raster(palette = "PuOr", 
            title = "Difference \nJSDM - Maxent \nNachtigall project",
            breaks = c(-1,-0.5,0, 0.5, 1),
            style = "cont")+
  tm_shape(water) +
  tm_polygons("lightblue", border.col = NULL)+
  tm_shape(berlin) +
  tm_borders("ivory4", lwd = 2) +
   tm_layout(main.title = "Jsdm - Maxent model - Nachtigall project 2018", 
            compass.type = "arrow",
            legend.title.size = 0.8,
            legend.text.size = 0.6, 
          legend.bg.color = "white",
          legend.position = c("right", "top"),
            inner.margins = c(0.05,0.05,0.05,0.1),
          legend.frame = TRUE)+
  tm_compass(position = c("right", "bottom"),
             size = 3) +
  tm_scale_bar(position=c("left", "bottom"), 
               size = 1) 

tmap_save(filename = paste0(results_wd, "/Comparison_JSDM-Maxent_nachtigall_map.png"), 
         width = 8, height = 7, units = "in")
```




